---
output:
  html_document:
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
---

Incubation
===

Masters experiment, incubation of soils amended with various amendments

## Load libraries

```{r, cache=TRUE, warning=FALSE, message=FALSE}
library(phyloseq)
library(tidyverse)
library(vegan)
library(ggpubr)
```

## Read in the data 
use `readRDS` to load phyloseq object
```{r, cache=TRUE}
inc.raw <- readRDS("Data/incubation_raw.RDS")
inc.raw
```

## Functions

```{r, cache=TRUE, warning=FALSE, message=FALSE}

# Put phyloseq object into a df with .02% phylum (glomed at phylum level)
RelativeAbundanceDf <- function(physeq) {
    physeq %>% tax_glom(taxrank = "Phylum") %>% transform_sample_counts(function(x) {
        x/sum(x)
    }) %>% psmelt() %>% filter(Abundance > 0.02) %>% arrange(Phylum)
}

# Function to plot relative abundance
PlotRelativeAbundance <- function(df) {
    ggplot(df, aes(x = as.factor(day), y = Abundance, fill = Phylum)) + 
    facet_grid(treatment ~ .) + 
    geom_bar(stat = "identity") +
    #scale_fill_manual(values = phylum.colors) + 
        # Remove x axis title
    theme(axis.title.x = element_blank()) + 
    guides(fill = guide_legend(reverse = TRUE, keywidth = 1, keyheight = 1)) + 
    ylab("Relative Abundance (Phyla > 2%) \n") +
    ggtitle("Phylum Composition of Incubation Soils \n Bacterial Communities by Treatment")
}

#Scale reads function to be used prior to ordination
ScaleReads <- function(physeq, n) {
  physeq.scale <- transform_sample_counts(physeq, function(x) {
    (n * x/sum(x))
  })
  otu_table(physeq.scale) <- floor(otu_table(physeq.scale))
  physeq.scale <- prune_taxa(taxa_sums(physeq.scale) > 0, physeq.scale)
  return(physeq.scale)
}

# Function to summarise a data frame and give statistics
DataSummary <- function(data, varname, groupnames) {
  require(plyr)
  SummaryFunc <- function(x, col) {
    c(mean = mean(x[[col]], na.rm = TRUE), sd = sd(x[[col]], na.rm = TRUE))
  }
  data.sum <- ddply(data, groupnames, .fun = SummaryFunc, varname)
  data.sum <- rename(data.sum, c(mean = varname))
}
```

Use function from above to create df with .02% of the phylum level of OTUs
```{r, cache=TRUE, warning=FALSE, message=FALSE, error=FALSE}
inc.raw.phylum.2percent <- RelativeAbundanceDf(inc.raw)

# Plot and save image
inc.phylum.abundance <- PlotRelativeAbundance(inc.raw.phylum.2percent)
tiff('Images/inc.phylum.abundance.tiff', units="in", width=5, height=5, res=300)
inc.phylum.abundance
dev.off()
inc.phylum.abundance
```

Some data wrangling to make the plots look nicer
```{r cache=TRUE, error=FALSE, message=FALSE, warning=FALSE}
# First split into two phyloseq objects
# Incubation
inc.treatment <- subset_samples(inc.raw, day %in% c("0", "7", "14", "21", "35", "49", "97"))
# Amends
inc.amend <- subset_samples(inc.raw, treatment %in% c("AlfalfaAmend", "CompostAmend"))

# Now let's pool the reps so that y-axis goes to 1, need to do for each object
inc.merged <- inc.treatment
variable.1 <- as.character(get_variable(inc.merged, "treatment"))
variable.2 <- as.character(get_variable(inc.merged, "day"))
sample_data(inc.merged)$TreatmentAndDay <- mapply(paste0, variable.1, variable.2, collapse = "-")
inc.merged <- merge_samples(inc.merged, "TreatmentAndDay")
sample_data(inc.merged)$treatment <- levels(sample_data(inc.treatment)$treatment)
```

## Relative abundance of phyla
```{r cache=TRUE, error=FALSE, message=FALSE, warning=FALSE}
# Innie to outtie, make df and plot relative abumndace
treatments.abundance <- PlotRelativeAbundance(RelativeAbundanceDf(inc.treatment))
amendments.abundance <- PlotRelativeAbundance(RelativeAbundanceDf(inc.amend))
mergeddf.abundance <- PlotRelativeAbundance(RelativeAbundanceDf(inc.merged))

# Save as images, high quality
tiff('Images/treatments.abundance.tiff', units="in", width=5, height=5, res=300)
treatments.abundance
dev.off()
tiff('Images/amendments.abundance.tiff', units="in", width=5, height=5, res=300)
amendments.abundance
dev.off()
tiff('Images/mergeddf.abundance.tiff', units="in", width=5, height=5, res=300)
mergeddf.abundance
dev.off()
treatments.abundance
amendments.abundance
mergeddf.abundance
```

## PCoA
We already have the function `ScaleReads` to re-sample our data to a specific read number per sample. 
```{r cache=TRUE, error=FALSE, message=FALSE, warning=FALSE}
# Now call the function with the phyloseq object you wish to scale
inc.scale.treatment <- ScaleReads(inc.treatment, n = 6000)
```
Fix day levels in `sample_data`
```{r cache=TRUE, error=FALSE, message=FALSE, warning=FALSE}
sample_data(inc.scale.treatment)$day <- factor(sample_data(inc.scale.treatment)$day, 
  levels = c("0", "7", "14", "21", "35", "49", "97"))
```
Now use the `ordinate` function from phyloseq
```{r cache=TRUE, error=FALSE, message=FALSE, warning=FALSE}
inc.ordination.treatment <- ordinate(physeq = inc.scale.treatment, method = "PCoA", distance = "bray")
```
Use `plot_ordination` to create the plot then manipulate it with ggplot2
```{r cache=TRUE, error=FALSE, message=FALSE, warning=FALSE}
inc.ordination.plot.treatment <- plot_ordination(physeq = inc.scale.treatment, ordination = inc.ordination.treatment, 
  color = "day", shape = "treatment", title = "PCoa of Incubation Bacterial Communities") + 
  #scale_color_manual(values = phylum.colors) + 
  geom_point(aes(color = day), alpha = 0.7, size = 4) + 
  geom_point(color = "grey90", size = 1.5)  

tiff('Images/inc.ordination.plot.treatment.tiff', units="in", width=10, height=10, res=300)
inc.ordination.plot.treatment
dev.off()

inc.ordination.plot.treatment
```

## Alpha Diversity
First remove OTUs that sum to 0
```{r cache=TRUE, error=FALSE, message=FALSE, warning=FALSE}
inc.treatment.pruned <- prune_species(speciesSums(inc.treatment) > 0, inc.treatment)
richness.data <- plot_richness(inc.treatment.pruned, measures = "Shannon")
richness.df <- richness.data$data

scaled.richness.data <- plot_richness(inc.scale.treatment, measures = "Shannon")
scaled.df <- scaled.richness.data$data

richness.shannon <- ggplot(richness.df, aes(x = day, y = value), color = treatment) +
  geom_boxplot(aes(color = treatment), position = "dodge") +
  facet_grid(~treatment)
richness.shannon
scaled.richness.shannon <- ggplot(scaled.df, aes(x = day, y = value), color = treatment) +
  geom_boxplot(aes(color = treatment), position = "dodge") +
  facet_grid(~treatment)
scaled.richness.shannon
```

## Statistics
Interesting results shown so far, we can see that the communities are changing over time and in response to nutrients
```{r cache=TRUE, error=FALSE, message=FALSE, warning=FALSE}
# Stats
# Adonis
inc.scale.df <- as(sample_data(inc.scale.treatment), "data.frame")
inc.distance <- distance(inc.scale.treatment, "bray")
inc.adonis <- adonis(inc.distance ~ treatment + day, inc.scale.df)
inc.adonis
# I'm pretty sure the order you feed the categories
incadonis.reverse <- (adonis(inc.distance ~ day + treatment, inc.scale.df))
```

## Soil chemical measurements
Plot the other data, mainly we want to explore how the nutrient concentration and microbial biomass numbers are changing during the course of the incubation.
```{r cache=TRUE, error=FALSE, message=FALSE, warning=FALSE}
inc.treatment
sample_data(inc.treatment)$day <- as.factor(sample_data(inc.treatment)$day)
inc.data <- as.data.frame(sample_data(inc.treatment))
colnames(inc.data)
```

## Nitrate "NO3"
```{r cache=TRUE, error=FALSE, message=FALSE, warning=FALSE}
inc.nitrate.eb <- DataSummary(inc.data, varname = "NO3", groupnames = c("day", "treatment"))
inc.nitrate.eb

plot.inc.nitrate.eb <- ggplot(inc.nitrate.eb, aes(x = day, y = NO3, group = treatment, color = treatment)) + 
  geom_errorbar(aes(ymin = NO3 - sd, ymax = NO3 + sd), width = 1, position = position_dodge(0.05)) +
  geom_line(aes(linetype = treatment)) + 
  geom_point(aes(shape = treatment)) + 
  labs(title = "Plot of NO3 by day", x = "Day", y = "NO3") + 
  theme_classic()
tiff('Images/nitrate.eb.tiff', units="in", width=5, height=5, res=300)
plot.inc.nitrate.eb
dev.off()

inc.nitrate <- ggplot(inc.data, aes(x = day, y = NO3, group = treatment)) + 
  geom_point(aes(color = treatment))
tiff('Images/nitrate.tiff', units="in", width=5, height=5, res=300)
inc.nitrate
dev.off()

plot <- ggplot(inc.data, aes(x = treatment, y = NO3, color = treatment)) +
  facet_grid(~day) +
  geom_boxplot(position = "dodge") +
  theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank())
tiff("Images/boxplot.nitrate.tiff", units = "in", width = 5, height = 5, res = 300)
plot
dev.off()

inc.data.7 <- inc.data %>%
  filter(day == 7) %>%
  ggplot(aes(x = treatment, y = NO3, color = treatment)) +
  geom_boxplot() +
  #facet_grid( ~ day) +
  rotate_x_text(angle = 45) +
  geom_hline(yintercept = mean(inc.data$NO3), color = "red") +
  stat_compare_means(method = "anova") +
  stat_compare_means(label = "p.signif", method = "t.test", ref.group = ".all.")
tiff("Images/nitrate.day.7.boxplot.tiff", units = "in", width = 5, height = 5, res = 300)
inc.data.7
dev.off()

test <- inc.data %>%
  filter(day == 7)

stat.day.7 <- ggboxplot(test, x = "treatment", y = "NO3", color = "treatment", legend = "none") +
  rotate_x_text(angle = 45) +
  geom_hline(yintercept = mean(test$NO3)) +
  stat_compare_means(method = "anova") +
  stat_compare_means(label = "p.signif", method = "t.test", ref.group = ".all.")
tiff('Images/stat.day.7.tiff', units="in", width=5, height=5, res=300)
#insert ggplot code
stat.day.7
dev.off()
plot.inc.nitrate.eb
inc.nitrate
plot
inc.data.7
stat.day.7
```

## Amonia "NH3" 
```{r cache=TRUE, error=FALSE, message=FALSE, warning=FALSE}
inc.amonia.eb <- DataSummary(inc.data, varname = "NH3", groupnames = c("day", "treatment"))

plot.inc.amonia.eb <- ggplot(inc.amonia.eb, aes(x = day, y = NH3, group = treatment, color = treatment)) +
  geom_errorbar(aes(ymin = NH3 - sd, ymax = NH3 + sd), width = 1, position = position_dodge(0.05)) + 
  geom_line(aes(linetype = treatment)) + 
  geom_point(aes(shape = treatment)) + labs(title = "Plot of NH3 by day", x = "Day", y = "NH3") + 
  theme_classic()
plot.inc.amonia.eb

inc.amonia <- ggplot(inc.data, aes(x = day, y = NH3, group = treatment)) + geom_point(aes(color = treatment))
inc.amonia
tiff('Images/inc.amonia.tiff', units="in", width=5, height=5, res=300)
#insert ggplot code
inc.amonia
dev.off()
tiff('Images/plot.inc.amonia.eb.tiff', units="in", width=5, height=5, res=300)
#insert ggplot code
plot.inc.amonia.eb
dev.off()
```

## Microbial Biomass 

```{r cache=TRUE, error=FALSE, message=FALSE, warning=FALSE}
inc.micr.biomass.c.eb <- DataSummary(inc.data, varname = "MBC_mg.kg_per_dry_wt_soil", groupnames = c("day", "treatment"))

plot.inc.micr.biomass.c.eb <- ggplot(inc.micr.biomass.c.eb, aes(x = day, y = MBC_mg.kg_per_dry_wt_soil, group = treatment, color = treatment)) + 
  geom_errorbar(aes(ymin = MBC_mg.kg_per_dry_wt_soil - sd, ymax = MBC_mg.kg_per_dry_wt_soil + sd), width = 1, position = position_dodge(0.05)) + 
  geom_line(aes(linetype = treatment)) + 
  geom_point(aes(shape = treatment)) + 
  labs(title = "Plot of MBC by day", x = "Day", y = "MBC") + 
  theme_classic()
plot.inc.micr.biomass.c.eb

inc.micr.biomass.c <- ggplot(inc.data, aes(x = day, y = MBC_mg.kg_per_dry_wt_soil, group = treatment)) + geom_point(aes(color = treatment))
inc.micr.biomass.c
tiff('Images/plot.inc.micr.biomass.c.eb.tiff', units="in", width=5, height=5, res=300)
plot.inc.micr.biomass.c.eb
dev.off()
tiff('Images/inc.micr.biomass.c.tiff', units="in", width=5, height=5, res=300)
inc.micr.biomass.c
dev.off()
```
## pH
```{r cache=TRUE, error=FALSE, message=FALSE, warning=FALSE}
inc.pH.eb <- DataSummary(inc.data, varname = "pH", groupnames = c("day", "treatment"))

plot.inc.pH.eb <- ggplot(inc.pH.eb, aes(x = day, y = pH, group = treatment, color = treatment)) + 
  geom_errorbar(aes(ymin = pH - sd, ymax = pH + sd), width = 1, position = position_dodge(0.05)) + 
  geom_line(aes(linetype = treatment)) + 
  geom_point(aes(shape = treatment)) + 
  labs(title = "Plot of pH by day", x = "Day", y = "MBC") + 
  theme_classic()
plot.inc.pH.eb

tiff('Images/plot.inc.pH.eb.tiff', units="in", width=5, height=5, res=300)
plot.inc.pH.eb
dev.off()

inc.pH <- ggplot(inc.data, aes(x = day, y = pH, group = treatment)) + geom_point(aes(color = treatment))
inc.pH

tiff('Images/inc.pH.tiff', units="in", width=5, height=5, res=300)
inc.pH
dev.off()
```
## Total Nitrogen via combustion analysis
```{r cache=TRUE, error=FALSE, message=FALSE, warning=FALSE}
inc.N_flash.eb <- DataSummary(inc.data, varname = "N_flash", groupnames = c("day", "treatment"))

plot.inc.N_flash.eb <- ggplot(inc.N_flash.eb, aes(x = day, y = N_flash, group = treatment, color = treatment)) + 
  geom_errorbar(aes(ymin = N_flash - sd, ymax = N_flash + sd), width = 1, position = position_dodge(0.05)) + 
  geom_line(aes(linetype = treatment)) + 
  geom_point(aes(shape = treatment)) + 
  labs(title = "Plot of N_flash by day", x = "Day", y = "% N") + 
  theme_classic()
plot.inc.N_flash.eb

tiff('Images/plot.inc.N_flash.eb.tiff', units="in", width=5, height=5, res=300)
plot.inc.N_flash.eb
dev.off()

inc.N_flash <- ggplot(inc.data, aes(x = day, y = N_flash, group = treatment)) + geom_point(aes(color = treatment))
inc.N_flash

tiff('Images/inc.N_flash.tiff', units="in", width=5, height=5, res=300)
inc.N_flash
dev.off()
```
## Total Carbon via combustion analysis
```{r cache=TRUE, error=FALSE, message=FALSE, warning=FALSE}
inc.C_flash.eb <- DataSummary(inc.data, varname = "C_flash", groupnames = c("day", "treatment"))

plot.inc.C_flash.eb <- ggplot(inc.C_flash.eb, aes(x = day, y = C_flash, group = treatment, color = treatment)) + 
  geom_errorbar(aes(ymin = C_flash - sd, ymax = C_flash + sd), width = 1, position = position_dodge(0.05)) + 
  geom_line(aes(linetype = treatment)) + 
  geom_point(aes(shape = treatment)) + 
  labs(title = "Plot of C_flash by day", x = "Day", y = "% C") + 
  theme_classic()
plot.inc.C_flash.eb

tiff('Images/inc.C_flash.tiff', units="in", width=5, height=5, res=300)
plot.inc.C_flash.eb
dev.off()

inc.C_flash <- ggplot(inc.data, aes(x = day, y = C_flash, group = treatment)) + geom_point(aes(color = treatment))
inc.C_flash

tiff('Images/inc.C_flash.tiff', units="in", width=5, height=5, res=300)
inc.C_flash
dev.off()
```
## Gravimetric moisture content
```{r cache=TRUE, error=FALSE, message=FALSE, warning=FALSE}
inc.gravimetric_water_content.eb <- DataSummary(inc.data, varname = "gravimetric_water_content", groupnames = c("day", "treatment"))

plot.inc.gravimetric_water_content.eb <- ggplot(inc.gravimetric_water_content.eb, aes(x = day, y = gravimetric_water_content, group = treatment, color = treatment)) + 
  geom_errorbar(aes(ymin = gravimetric_water_content - sd, ymax = gravimetric_water_content + sd), width = 1, position = position_dodge(0.05)) + 
  geom_line(aes(linetype = treatment)) + 
  geom_point(aes(shape = treatment)) + 
  labs(title = "Plot of gravimetric_water_content by day", x = "Day", y = "gravimetric_water_content") + 
  theme_classic()
plot.inc.gravimetric_water_content.eb

tiff('Images/plot.inc.gravimetric_water_content.eb.tiff', units="in", width=5, height=5, res=300)
plot.inc.gravimetric_water_content.eb
dev.off()

inc.gravimetric_water_content <- ggplot(inc.data, aes(x = day, y = gravimetric_water_content, group = treatment)) + geom_point(aes(color = treatment))
inc.gravimetric_water_content

tiff('Images/inc.gravimetric_water_content.tiff', units="in", width=5, height=5, res=300)
inc.gravimetric_water_content
dev.off()
```

## Under Construction
Try tax_glom to identify the OTUs on one group but not another

```{r cache=TRUE, error=FALSE, message=FALSE, warning=FALSE}
inc.raw.control <- subset_samples(inc.raw, treatment == "Control" & day == "0")
inc.raw.alfalfa <- subset_samples(inc.raw, treatment == "Alfalfa" & day == "0")
# Day zero comparison of control and alfalfa OTUs
control.no.0 <- filter_taxa(inc.raw.control, function(x) sum(x) >0, TRUE)
alfalfa.no.0 <- filter_taxa(inc.raw.alfalfa, function(x) sum(x) >0, TRUE)
control.taxa <- rownames(tax_table(control.no.0))
alfalfa.taxa <- rownames(tax_table(alfalfa.no.0))
length(intersect(control.taxa, alfalfa.taxa))
# OTUs in alfalfa day 0 only 
only.alfalfa <- setdiff(alfalfa.taxa, control.taxa)
length(only.alfalfa)
tax.in.alf <- tax_table(inc.raw.alfalfa)[only.alfalfa]
# Below melt for plotting and prune to get taxa from larger phyloseq object
#only.alfalfa.day.0 <- prune_taxa(only.alfalfa, inc.raw.alfalfa)
#only.alfalfa.day.0.df <- psmelt(only.alfalfa.day.0)
```







































